---
title: "Workshopping project proposals"
subtitle: "Lecture 6"
date: "May 23, 2023"
format: revealjs
---

## Today

-   Fit between research questions and data
-   Workshop your research questions

## Fit between research questions and data

::: incremental
-   Is the sample generalizable to the target population?
-   Does it contain the necessary variables to answer the research question?
:::

## Research questions and variables {.smaller}

-   EADA data

```{r, include = FALSE}
library(tidyverse)
sports <- readRDS("../exampleprep/sports_npeople.rds") |> 
  select(-player_coach_ratio)
```

```{r}
glimpse(sports)
```

## Would these research questions work? What variables would they use? {.smaller}

-   How does the number of assistant coaches assigned to teams vary by university?

```{r}
glimpse(sports)
```

## Would these research questions work? What variables would they use? {.smaller}

-   How does the number of coaches assigned to basketball teams vary by gender of the team?

```{r}
glimpse(sports)
```

## Would these research questions work? What variables would they use? {.smaller}

-   How is the number of games a team wins related to the number of coaches it has?

```{r}
glimpse(sports)
```

## Workshopping {.smaller}

-   Read your group members' submissions

-   Answer some questions about their work

    -   Return feedback in github: copy/paste template from project instructions; fill out
    -   10-15 min/person
    -   Goal is improvement--how can they make their work better?

-   Talk through your thoughts on each group member's proposal and their choice of question (5-10 minutes each)

------------------------------------------------------------------------

-   Was this helpful?
-   Do you feel like you're on a good path with your project?
-   Questions you'd like me to address?
