---
title: "Describing data: part 2"
subtitle: "Lecture 8"
date: "May 31, 2023"
format: revealjs
---

```{r, include = FALSE}
library(tidyverse)
library(openintro)
sports <- readRDS("data/sports_teams.rds")
```

## Logistics {.smaller}

-   Project proposals

    -   I will get my feedback to you soon--will be posted as a github issue (grades will be in Sakai)
    
. . . 

-   Project descriptive statistics

    -   Due **Tuesday June 6**. Instructions and example are posted on the website.
    
. . . 

-   Homework 1

    -   Cancelled as a mandatory assignment; time is short
    -   BUT practice is the only way to learn this stuff!
    -   So, extra credit opportunities will be worked into exercises--I highly recommend doing as much as you can!

## Today

-   Finish univariate summaries
-   Preparing data frames

# Where we left off: summarizing numeric variables

## Summarizing a distribution {.smaller}

::: columns
::: {.column width="50%"}
::: incremental
-   Center

    -   median(dataset\$var, na.rm = TRUE)
    -   mean(dataset\$var, na.rm = TRUE)

-   Spread

    -   quantile(dataset\$var, na.rm = TRUE)
    -   sd(dataset\$var, na.rm = TRUE)

-   (Almost) everything

    -   summary(dataset\$var)
:::
:::

::: {.column width="50%"}
```{r, echo = FALSE, fig.height=4, fig.width=5}
normdata <- tibble(d = rnorm(1000, mean = 5, sd = 1))
ggplot(normdata, aes(x = d)) + 
  geom_histogram(binwidth = .25) + 
  theme_minimal() + 
  labs(x = "", y = "") + 
  theme(axis.text.y = element_blank())
```
:::
:::

## Exercise question 2 {.smaller}

```{r}
glimpse(acs12)
```

## Question 2 solutions {.smaller}

What is the mean of `hrs_work`?
What is the median?
(there are two ways to get this info)

::: fragment
Approach 1:

```{r}
mean(acs12$hrs_work, na.rm = TRUE)
median(acs12$hrs_work, na.rm = TRUE)
```
:::

::: fragment
Approach 2:

```{r}
summary(acs12$hrs_work)
```
:::

## Question 2 solutions {.smaller}

What is the standard deviation of `hrs_work`?

::: fragment
```{r}
sd(acs12$hrs_work, na.rm = TRUE)
```
:::

## Question 2 solutions {.smaller}

What proportion of people in the data set are missing information (have NAs) for this variable?
(there are many ways to do this)

::: fragment
```{r}
# Find the number of NA values
summary(acs12$hrs_work)
```

```{r}
# Find the total number of observations
glimpse(acs12)
```
:::

## Question 2 solutions

::: fragment
Now that we have those values, we can use R to do the math

```{r}
# Divide!
1041/2000
```
:::

## Shape

::: columns
::: {.column width="50%"}
```{r, echo = FALSE, fig.height=4, fig.width=5}
normdata <- tibble(d = rnorm(1000, mean = 5, sd = 1))
ggplot(normdata, aes(x = d)) + 
  geom_histogram(binwidth = .25) + 
  theme_minimal() + 
  labs(x = "", y = "") + 
  theme(axis.text.y = element_blank())
```
:::

::: {.column width="50%"}
```{r, echo = FALSE, fig.height=4, fig.width=5}
unifdata <- tibble(d = runif(1000, min = 1.5, max = 8.5))
ggplot(unifdata, aes(x = d)) + 
  geom_histogram(binwidth = .25) + 
  theme_minimal() + 
  labs(x = "", y = "") + 
  theme(axis.text.y = element_blank())
```
:::
:::

## Shape

-   We'll go over how to make plots later
-   But summary statistics tell you things about shape too

## Skew

::: columns
::: {.column width="50%"}
```{r, echo = FALSE, fig.height=4, fig.width=5}
rightdata <- tibble(d = rbeta(1000, 1, 10))
ggplot(rightdata, aes(x = d)) + 
  geom_histogram(bins = 30) + 
  theme_minimal() + 
  labs(x = "", y = "") + 
  theme(axis.text.y = element_blank(),
        axis.text.x = element_blank())
```

Right-skewed data
:::

::: {.column width="50%"}
```{r, echo = FALSE, fig.height=4, fig.width=5}
leftdata <- tibble(d = rbeta(1000, 10, 1))
ggplot(leftdata, aes(x = d)) + 
  geom_histogram(bins = 30) + 
  theme_minimal() + 
  labs(x = "", y = "") + 
  theme(axis.text.y = element_blank(),
        axis.text.x = element_blank())
```

Left-skewed data
:::
:::

## Skew

-   Extreme values influence the mean more than the median
-   When mean is higher than median: data might be skewed right (and vice versa)

## Example: American Community Survey commute time

time_to_work: Travel time to work, in minutes.

```{r}
mean_commute <- mean(acs12$time_to_work, na.rm = TRUE)
median_commute <- median(acs12$time_to_work, na.rm = TRUE)
```

::: fragment
```{r}
mean_commute
```
:::

::: fragment
```{r}
median_commute
```
:::

## Example: American Community Survey commute time

```{r, echo = FALSE, warning = FALSE}
ggplot(acs12, aes(x = time_to_work)) + 
  geom_histogram(bins = 20) + 
  geom_vline(xintercept = mean_commute, color = "blue") +
  geom_text(color = "blue", aes(x = mean_commute + 8, y = 242, label = "mean")) + 
  geom_vline(xintercept = median_commute, color = "red") + 
  geom_text(color = "red", aes(x = median_commute - 8, y = 242, label = "median")) + 
  theme_minimal() +
  labs(x = "Commute time (minutes)", y = "Number of observations")
```

## Shape: Quartiles

-   Location of quartiles also tells you something about shape

```{r}
summary(acs12$time_to_work)
```

## Review {.smaller}

-   Categorical data: look at the available categories and how many observations are in each category

    -   `unique(dataframe$variable)`
    -   `table(dataframe$variable, useNA = "always")`
    -   `table(dataframe$variable1, dataframe$variable2, useNA = "always")`

## Review {.smaller}

-   Numeric data: Look at summary statistics that tell you something about the distribution
    -   `summary(dataframe$variable)`
    -   `sd(dataframe$variable, na.rm = TRUE)`

# Modifying data frames

## Removing observations

::: incremental
-   You may not be interested in all observations

-   Example:

    -   RQ: How is commute time related to hours worked?
:::

## Removing observations {.smaller}

-   How is commute time related to hours worked?
-   Data: ACS 2012

::: fragment
```{r}
glimpse(acs12)
```
:::

::: fragment
```         
-   Need to remove people who are unemployed/not in labor force (and don't have a job or a commute!)
```
:::

## Removing observations {.smaller}

::: incremental
-   How? With `filter()`
-   `filter(dataframe, condition)`
:::

::: fragment
```{r}
acs12_employedonly <- filter(acs12, employment == "employed")
```
:::

::: fragment
-   Did it work? Check with `table()`

```{r}
table(acs12_employedonly$employment, useNA = "always")
```
:::

## Conditions {.smaller}

::: incremental
-   `filter(acs12, employment == "employed")` 
-   condition: `employment == "employed"`
-   This specifies which observations you want to *keep* 
:::

::: fragment
-   Common comparison operators:
    -   `==` : equal to (note there are two equals signs!)
    -   `!=` : not equal to
    -   `>`, `>=` : greater than, greater than or equal to
    -   `<`, `<=` : less than, less than or equal to
:::

::: fragment
-   Values:

    -   Can be numbers, letters/words, or TRUE/FALSE--should match the response options of your variable
    -   Put letters/words in quotation marks
:::

## Condition examples {.smaller}

::: incremental
-   `filter(acs12, citizen == "no")`
-   `filter(acs12, income <= 12000)`
-   `filter(acs12, birth_qrtr != "jan thru mar")`
-   `filter(acs12, hrs_work > 20)`
:::

## Other useful condition operators {.smaller}

::: incremental
-   Two or more requirements

    -   `&` : and

    -   `|` : or

    -   `filter(acs12, citizen == "no" & lang == "english")`

    -   `filter(acs12, race == "black" | race == "asian")`

-   Missing values

    -   `is.na()`

    -   `!is.na()`

        -   is *not* missing -- good for removing rows with missing values

    -   `filter(acs12, !is.na(income))`

-   More: [Click here for R for Data Science](https://r4ds.had.co.nz/transform.html#filter-rows-with-filter)
:::

## Common mistakes and error messages {.smaller}

```{r, error = TRUE}
filter(acs12, employment = "employed")
```

## Common mistakes and error messages {.smaller}

```{r, error = TRUE}
filter(acs12, employment == employed)
```

## Common mistakes and error messages {.smaller}

```{r, error = TRUE}
filter(acs12, employment == "employd")
```

## Common mistakes and error messages {.smaller}

```{r, error = TRUE}
filter(acs12, employment == "Employed")
```

## Exercise: Filtering

-   Clone and open the project repo now (**ex-5-31-yourusername**)

-   Then open the .qmd file and try out some filtering

    -   [Instructions: Computing -\> Cloning and committing](https://soc333-sum23.github.io/computing-clonerepo.html)
