---
title: "Describing data: part 3"
subtitle: "Lecture 9"
date: "June 1, 2023"
format: revealjs
---

```{r, include = FALSE}
library(tidyverse)
library(openintro)
sports <- readRDS("data/sports_teams.rds")
```

## Logistics {.smaller}

-   Project proposal

    -   All materials are posted
    -   Due Tuesday June 6
    -   We will save some time to work on it in class Monday June 5

## Today

-   Creating new variables
-   Plots

## Creating variables

::: incremental
-   You may want to create new variables/change the response options of old variables

-   Example:

    -   RQ: How does employment status vary between teenagers, adults below retirement age (67), and adults above retirement age?
:::

## Creating variables {.smaller}

-   How does employment status vary by age category?

```{r}
glimpse(acs12)
```

::: incremental
-   We need to take the "age" variable and use it to create age categories (making a categorical variable from a numeric variable!)
:::

## Exercise: Creating new variables

-   Clone and open the project repo now (**ex-5-31-yourusername**)
-   Then open the .qmd file and try out some filtering

    -   [Instructions: Computing -\> Cloning and committing](https://soc333-sum23.github.io/computing-clonerepo.html)

