---
title: "Describing data: part 4"
subtitle: "Lecture 10"
date: "June 5, 2023"
format: revealjs
smaller: true
---

```{r, include = FALSE}
library(tidyverse)
library(openintro)
```

## Logistics {.smaller}

-   Project component 2: descriptive statistics

    -   All materials are posted (instructions, example, github repos)
    -   Due **Wednesday June 7** 11:59pm
    -   I have returned feedback (GitHub issue on your repo) and grades (Sakai) for your proposals
    -   We have covered everything for parts 1-3. Part 4 is today and tomorrow.
    -   I'm around before/after class for questions (after generally better), or set up an appointment

## Today {.smaller}

-   Answers to filtering
-   Pipes
-   Working on creating variables
-   Answers to creating variables
-   Plots

## Filtering exercise (ex-5-31): answers! {.smaller}

-   Find this on your computer (no need to clone it again)

## Stringing commands together with pipes (`|>`)

::: incremental
-   Often we need to change data frames in more than one way
-   Example from last week: How does employment status vary by age category?
-   We need to create an age category variable (as we talked about Thursday)
-   But we probably want to filter too--why analyze kids?
:::

## Stringing commands together: approach 1

We can do this with the tools we have.

. . .

```{r}
# first we make the new variable
acs12_newagevar <- mutate(acs12, agecat = case_when(age < 14 ~ "child",
                                                    age < 18 ~ "teen",
                                                    age < 67 ~ "adult",
                                                    TRUE ~ "retired"))

# then we filter to remove children
acs12_nokids <- filter(acs12_newagevar, agecat != "child")

# did it work?
table(acs12_nokids$agecat, useNA = "always")
```

. . .

-   But it's kind of ugly... we don't need to save that middle step

## Stringing commands together: approach 2, with `|>`

::: incremental
-   The pipe operator, `|>`, lets us pass the result of one function directly into another one
-   The `|` symbol is the key below "delete" on your keyboard (not I, not l, not 1)
-   It replaces the first mutate/filter argument (the dataset)
-   "Take the thing that came before this and give it to the function that comes after this"
:::

. . .

```{r}
acs12_nokids <- acs12 |> # start with acs12
  mutate(agecat = case_when(age < 14 ~ "child", #then add a new variable to it
                            age < 18 ~ "teen",
                            age < 67 ~ "adult",
                            TRUE ~ "retired")) |> 
  filter(agecat != "child") # then filter out the kids

table(acs12_nokids$agecat, useNA = "always")
```

## Another `|>` example

-   This:

```{r, eval = FALSE}
mutate(acs12, agecat = case_when(age < 14 ~ "child",
                                 age < 18 ~ "teen",
                                 age < 67 ~ "adult",
                                 TRUE ~ "retired"))
```

-   Is the same as this:

```{r, eval = FALSE}
acs12 |> 
  mutate(agecat = case_when(age < 14 ~ "child",
                            age < 18 ~ "teen",
                            age < 67 ~ "adult",
                            TRUE ~ "retired"))
```

## More on `|>`

::: incremental
-   We're just scratching the surface in this class

-   When you need to clean your data or when your analyses are more complex, `|>` makes your life a lot easier!

-   Sometimes in internet resources more than \~2 years old, you'll see `%>%` instead--this older version does exactly the same thing

-   Annoying to type out?
    There's a keyboard shortcut

    -   on mac, command-shift-M
    -   on pc, ctrl-shift-M
:::

## Continuing on creating new variables {.smaller}

-   Find your exercise from last Thursday (ex-6-01)

    -   No need to clone it again; find where it's saved on your computer

-   There are hints on the website--use them if you get stuck

-   10-15 minutes to work

# Creating new variables: answers!
