---
title: "Interpreting results"
subtitle: "Lecture 17"
date: "June 15, 2023"
format: 
  revealjs:
    smaller: true
---

```{r, include = FALSE}
library(tidyverse)
library(openintro)
library(infer)
library(patchwork)
set.seed(2)
```

## Logistics

-   Four classes left. No class Monday. We're almost there! :)

## Wrapping up the project

-   Project component 3: submit for grading by next Tuesday (June 20)

. . .

-   Final paper

    -   use (revised!) elements of your proposal, descriptive stats, and results to turn your question and analysis into a (concise) paper.
    -   More detailed instructions and example will be posted soon
    -   Due Wednesday June 28 11:59pm

. . .

-   Final presentation

    -   Short! \~5-10 minutes
    -   Briefly introduce your question, say a bit about why it's interesting, show your results, and talk about what you conclude from them.
    -   Most on Monday June 26 (last day of class)

## Final presentation logistics

-   **Looking for a few volunteers to go on Thursday June 22 instead!**

    -   Benefits...
    -   More time afterwards to focus on your final paper (finish up the semester early?)
    -   No stress while you watch day 2
    -   Everyone gets more flexibility on time
    -   What's your favorite kind of donut? :)

## Today {.smaller}

-   Interpreting results!

## Yesterday's exercise

## The (approximate) data analysis process {.smaller}

-   Determine topic ✓
-   Find data; learn what observations and variables are available ✓
-   Write research question ✓
-   Describe distributions of relevant variables ✓
-   Prepare data frame for analysis ✓
-   Describe relationships between variables ✓
-   Perform statistical tests ✓
-   **Communicate results**

## So you have a p value...

-   Rejecting/failing to reject hypotheses is useful
-   But it's also statistics-speak, not effective communication

## What do you want to be able to say?

-   Something about what the world is like!
-   Translate back from hypothesis test language---zoom back out. If you were describing your results to a friend, what would you say you learned about the world?

## What *can't* we say?

::: incremental
-   Knowing the limitations of your conclusions is super important!
-   Limitations can arise from several places.
:::

## Sampling and generalizability

::: incremental
-   Who can your results speak to?
-   Is your sample big enough?
-   Was it selected randomly?
:::

## Accuracy

::: incremental
-   Are people likely to give you accurate answers?

-   Unintentional inaccuracies

    -   Desirability biases
    -   Memory problems

-   Misrepresenting/lying

    -   Is your data coming from something or someone who has a stake in it?

-   Saying vs doing

    -   What people say is often different than what they do
    -   Stated ideals often don't match up with actions
    -   "I believe in equality but I oppose all the policies that would create it"
    -   Did your data collect information on saying, or on doing?
    -   Avoid drawing conclusions about what people would do when the data you have is about what they say
:::

## Operationalization

::: incremental
-   Are the questions and categories what you would like them to be?

-   Do we lose information through bad category choice or bad documentation?

    -   For example: bad but common operationalizations

        -   Race: white or nonwhite
        -   Gender: male or female
:::

## Causality

-   Association vs causation: What's the difference?

. . .

-   Association: two things are related
-   Causation: one thing drives another thing

. . .

-   It's nice to be able to say something about causation, but it's hard!
-   Hypothesis tests speak to association. They *sometimes* speak to causation as well---it all depends on the set up of your study!

## Exercise Q1

-   Does this provide evidence that using the internet helps people live longer? Why or why not?
-   Write down a plausible explanation for this relationship.

```{r, echo = FALSE}
facts <- cia_factbook |> 
  filter(!is.na(internet_users), 
         !is.na(life_exp_at_birth)) |> 
  mutate(percinternet = internet_users/population*100)

ggplot(facts, 
       aes(x = percinternet, y = life_exp_at_birth)) + 
  geom_point() + 
  theme_minimal() + 
  labs(title = "Life expectancy vs percent internet users in countries",
       subtitle = "2014 CIA factbook",
       x = "Percent internet users",
       y = "Life expectancy at birth (years)")
```

```{r, eval = FALSE, echo = FALSE}
# 1. Calculate test statistic
test_stat <- facts |> 
  specify(explanatory = percinternet,
          response = life_exp_at_birth) |> 
  calculate(stat = "slope")

# 2. Simulate null distribution
null_dist <- facts |> 
  specify(explanatory = percinternet,
          response = life_exp_at_birth) |> 
  hypothesize(null = "independence") |>
  generate(reps = 100000) |>
  calculate(stat = "slope")

# 3. Use both to get the p value
get_p_value(null_dist,
            obs_stat = test_stat,
            direction = "two-sided")
```

## Reading for Tuesday

-   There's an article linked on the course calendar. Mess around with the interactive element!

## Workshopping results

-   Teams

-   Same as last time: you have access to each other's repos on GitHub. Leave your feedback as an issue. Questions to address are at the top of the instructions page on the website ([link](https://soc333-sum23.github.io/project-results.html)).
