---
title: "5-31-23 exercise: filtering"
format: html
editor: visual
---

## Setup

First, we need to load some packages.
Click the play button to run the code in the block below (remember, you only need to do this once per session).

```{r}
library(tidyverse)
library(openintro)
```

## Filtering

Like yesterday, we will be working with the `acs12` data set.

Often, we only want to analyze adult respondents.
In the chunk below, create a modified version of `acs12` that contains only respondents who are 18 or older.
Call your new data frame `acs_adults`.

```{r}
acs_adults <- filter(acs12, age >= 18)
```

Look at the summary statistics of the `age` variable in your new, filtered data frame.
Did the filtering work?

```{r}
summary(acs_adults$age)
# yes, filtering worked! Our minimum age is now 18
```

How many respondents are left in this new data frame?
(we started with 2000)

(hint: there are multiple ways to do this, some of which involve code, one of which involves looking at the entry for your new data frame in the "environment" RStudio panel)

```{r}
glimpse(acs_adults)
# glimpse shows us that there are 1561 rows in the data frame now. We can also see this by looking at the information on acs_adults in the environment tab.
```

## Extra credit

Use the chunks below to create other subsets of the `acs12` data.
Do as many or as few as you feel are useful to you.
When you feel like you've got the basics down, try out some of the more complex filtering operators we discussed.

Check your results by looking at summary statistics or tables.
Look at your new data frames by opening them up from the environment panel.

#### For practice with simple comparison operators (`==`, `!=`, `<`, `<=`, `>`, `>=`):

-   Include only US citizens

    ```{r}
    acs_citizens <- filter(acs12, citizen == "yes")

    table(acs_citizens$citizen)
    ```

-   Include only people who are not White

    ```{r}
    acs_poc <- filter(acs12, race != "white")

    table(acs_poc$race)
    ```

-   Include only people who are retirement age (\>67).

    ```{r}
    acs_retired <- filter(acs12, age > 67)

    summary(acs_retired$age)
    ```

-   Exclude people who were born in winter

    ```{r}
    glimpse(acs12) # variable is birth_qrtr
    unique(acs12$birth_qrtr) # value is "jan thru mar"

    acs_nowinter <- filter(acs12,
           birth_qrtr != "jan thru mar")

    table(acs_nowinter$birth_qrtr)
    ```

#### For practice with more complex comparison operators:

-   Include only the people who are not in the labor force but still say they work more than 20 hours per week (??)

    ```{r}
    unique(acs12$employment) # "not in labor force" is the value we're interested in

    acs_confusingwork <- filter(acs12, hrs_work > 20 & employment == "not in labor force")

    # table checks the employment variable--we only have people who are not in the labor force
    table(acs_confusingwork$employment)
    # and summary checks hrs_work: we only have people who work more than 20 hours.
    summary(acs_confusingwork$hrs_work)
    ```

-   Include only Black and Asian respondents

    ```{r}
    unique(acs12$race)

    acs_blackasian <- filter(acs12, 
                             race == "black" | race == "asian")

    table(acs_blackasian$race)
    ```

-   Include only the people who reported a usable (not-NA) commute time

    ```{r}
    acs_withcommutes <- filter(acs12, !is.na(time_to_work))

    # no NA's reported!
    summary(acs_withcommutes$time_to_work)
    ```

-   Include people who have a disability or are unemployed

    ```{r}
    # checking for variable names: we want "disability", which has yes/no values, and "employment", where the relevant value is "unemployed"
    glimpse(acs12)

    acs_disabled_or_unemployed <- filter(acs12,
                                         disability == "yes" | employment == "unemployed")

    # the table shows that we have people with disabilities, people who are unemployed, and people who both have a disability and are unemployed. But there are no people who neither have a disability nor are unemployed. So this worked as expected!
    table(acs_disabled_or_unemployed$disability, acs_disabled_or_unemployed$employment)

    ```

-   Include non-retired adults (people who are older than 18 but younger than 67)

    ```{r}
    # I should have phrased this question better; really if we're interested in adults it should be 18 or older, not older than 18.
    acs_workingadults <- filter(acs12, age >= 18 & age < 67)

    ```

#### Other subsets that interest you:

```{r}

```

## Submission

When you are done, commit your changes and push them to GitHub!
([instructions](https://soc333-sum23.github.io/computing-clonerepo.html))
