---
title: "Hypothesis testing"
format: html
editor: visual
---

# Exercise: Hypothesis testing

## Setup: run these two chunks first!

```{r, eval = FALSE}
# the infer package is new, so we need to install it.
# this one only needs to be run one time
install.packages("infer")
```

```{r setup}
library(tidyverse)
library(openintro)
# we're adding the infer package to our lineup today
library(infer)
```

## Q1: Determining the correct hypothesis test, part 1. *See slides from lecture 15 for solutions*

For each of these research questions, identify the explanatory variable, the response variable, their types, and the correct statistical test. *Think about what the data would look like for each individual in these cases*

a. How does the amount of money a country spends on healthcare (per person) affect its average life expectancy?

    -   Explanatory variable(s) and type(s)?
    -   Response variable and type?
    -   Correct statistical test?

## Q2: Determining the correct hypothesis test, part 2. *See slides from lecture 15 for solutions*

Same thing, new questions. For each, identify the explanatory variable, the response variable, their types, and the correct statistical test.

b. How is a patient's body mass index related to the probability a doctor refers them to a specialist?

    -   Explanatory variable(s) and type(s)?
    -   Response variable and type?
    -   Correct statistical test?

c. How do outcomes of traffic stops (which can be warnings, citations, or arrests) vary by the driver's race?

    -   Explanatory variable(s) and type(s)?
    -   Response variable and type?
    -   Correct statistical test?

## Q3: Running your first hypothesis test!

We will be returning to the `acs12` data for this exercise. We will be using two variables: `gender` and `married`.

Both are categorical variables with two categories (male/female for gender, yes/no for married), so we'll use a two-sample Z test, just like in the class example.

Research question: How is someone's gender identity related to their probability of being married?

Here's a table to show the relationship between the two variables.

```{r}
table(acs12$gender, acs12$married)
```

### Step 1: Calculate the test statistic

Then look at it in the data viewer. What is the Z score of this sample?

**The Z score of this sample is about 1.6**

(You might notice that R gives you a warning message. This is fine in this case! It just means we shouldn't read into the positive/negative sign on the Z score. But since we're doing a two-tailed test here, that didn't matter anyway.)

```{r}
test_stat <- acs12 |> 
  specify(explanatory = gender,
          response = married,
          success = "yes") |> 
  hypothesize(null = "independence") |> 
  calculate(stat = "z")
```

### Step 2: Simulate the null distribution

```{r}
# 1. copy your code from above
# 2. change the name of the object from test_stat to null_dist
# 3. add generate(reps = 1000) in the correct place, like in the class example

null_dist <- acs12 |> 
  specify(explanatory = gender,
          response = married,
          success = "yes") |> 
  hypothesize(null = "independence") |> 
  generate(reps = 1000) |> 
  calculate(stat = "z")
```

### Step 3: Calculate the p value of your sample

Then interpret it. At a threshold value of p = 0.05, do you reject or fail to reject the null hypothesis that about the same number of men are married as women?

**I fail to reject the null hypothesis that about the same number of men are married as women. My p value, 0.116, is above my cutoff value of 0.05.**

```{r}
get_p_value(null_dist,
            obs_stat = test_stat,
            direction = "two-sided")
```

### Step 4: Visualize where your sample is with respect to your null distribution

```{r}
visualize(null_dist) +
  shade_p_value(obs_stat = test_stat, direction = "two-sided")
```

# Extra Credit

## Q4: Run a hypothesis test to answer the research question below: 

How is gender related to citizenship status among people who live in the US? 

Use the `acs12` data. 

## a: What are your null and alternative hypotheses? 

Null: Gender is not related to citizenship status.
Alternative: Gender is related to citizenship status.

## b: What are the explanatory and response variables, and what types are they? What kind of test do you need to run?

Explanatory: gender; categorical with two categories
Response: citizen; categorical with two categories
Test: 2 sample Z test

```{r}
# use this chunk to help you get a look at your data!
glimpse(acs12)
table(acs12$citizen)
table(acs12$gender)
```

## c: Run your test!

```{r}
# first: calculate your test statistic
test_stat <- acs12 |> 
  specify(explanatory = gender, response = citizen, success = "yes") |> 
  hypothesize(null = "independence") |> 
  calculate(stat = "z")

# then: simulate your null distribution
null_dist <- acs12 |> 
  specify(explanatory = gender, response = citizen, success = "yes") |> 
  hypothesize(null = "independence") |> 
  generate(reps = 1000) |> 
  calculate(stat = "z")

# next: calculate your p value 
get_p_value(null_dist, obs_stat = test_stat, direction = "two-sided")

# finally: visualize where you sample is with respect to your null distribution
visualize(null_dist) +
  shade_p_value(obs_stat = test_stat, direction = "two-sided")
```

## d: Interpret your results

With a threshold value of p = .05, do you reject or fail to reject your null hypothesis? 

I fail to reject the null hypothesis that gender is related to citizenship status. There is not sufficient evidence in these data to conclude that there is a relationship. Women who live in the US are about as likely to be citizens as men. 